### 2015-09-21 report.tex
updates to report and forward-reduction diagram

\documentclass{article}
\usepackage{graphicx}
\usepackage{minted}
\usepackage{parskip}
\usepackage[colorlinks,
            urlcolor=blue]{hyperref}
%\usepackage[margin=1in]{geometry}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{fancyvrb}
\usepackage{booktabs}
\usepackage{url}
\usepackage{hyperref}
\usepackage{multirow}
\usepackage{pbox}
\usepackage[margin=1in]{geometry}

\author{Ashwin Srinath}
\title{An OpenCL-based tridiagonal solver for evaluation
    of compact finite differences on parallel architectures}
\begin{document}
% \renewcommand{\bibname}{References}

\section{Introduction}

    The motivation for this work comes from the need to evaluate
    \emph{compact finite differences} arising in large simulations of combustion.
    Compact finite different schemes result in
    high-order of accuracy for relatively small stencils.
    Numerical methods with small stencil widths are especially benefitial
    in managing communication costs in large-scale cluster computing.
    This work is concerned with implementing algorithms for evaluating
    compact finite differences to make full use of available parallel architecture.
    The target architecture studied is the NVIDIA Tesla K20 GPU.
    The algorithm design for this architecture is presented,
    along with the associated issues.
    Results are presented for domains sized up to $2048^3$ grid points,
    using up to 64 GPUs.

\section{Related work}

    [FastTridiagonal] discuss the GPU implementation of three algorithms
    and their hybrid variants: Cyclic Reduction, Parallel Cyclic Reduction
    and Recursive Doubling.
    Their approach utilizes the GPUs \emph{shared memory},
    and addresses the concerns of bank conflicts arising in these algorithms.
    A global memory implementation of their solvers
    runs about three times slower, but is able to accomodate larger systems.
    The speedups they report are 12x over a 2.5 GHz quad-core CPU.

    [ScalableTridiagonal] and [Autotuned] both propose hybrid
    Parallel Cyclic Reduction and thread-level parallel Thomas algorithm (p-Thomas), and
    methods to determine the switching point for the algorithms.
    The speedups reported are 8.3x and 11x over a 3.33 GHz Intel i7 quad-core
    processor and a 3.4 GHz Intel i5 dual-core respectively.

    In [ScalableNumericallyStable],
    the authors implement a heterogenous (GPU+MPI) solver using the
    SPIKE algorithm for GPUs with a diagonal pivoting strategy for stability.
    More recently, [ChunkedCyclic] propose a \emph{chunked} cyclic reduction
    strategy for GPUss
    In [GPUApplicationForHigherOrderCFD], the inverse of the coefficient matrix
    is used to solve tridiagonal systems resulting from high-order compact
    finite difference schemes. The work here is for a single GPU.
    [An efficient GPU implementation of cyclic reduction solver] demonstrate
    a method to improve memory accesses for Cyclic Reduction
    using a reordering strategy, and apply it to compressible viscous flow
    simulations.

    --- Mention the speedups

    In this work, we propose a variant of Cyclic Reduction for solving
    tridiagonal systems arising in compact finite differences
    in a cluster environment (several GPU nodes).
    We introduce a strategy to
    reduce global memory accesses and decrease memory requirements by
    pre-computing coefficients appearing in Cyclic Reduction.

\section{Numerical method}

    We consider the following fourth-order
    compact finite difference scheme for evaluating the first derivative,
    with a 3-point stencil:

    \begin{equation}
        \alpha(f^{\prime}_{i-1} + f^{\prime}_{i+1}) + f^{\prime}_i
        =
        a\frac{f_{i+1} - f_{i-1}}{dx}
    \end{equation}

    For the particular scheme used, $\alpha = 1/4$,
    and $ a = 3/4 $.
    It is easily noted that this leads to a tridiagonal system of equations,
    along with a 3-point stencil to evaluate the RHS.
    Further, the left boundary can be treated using the following implicit equation:

    \begin{equation}
        f^{\prime}_1 + 2f^{\prime}_2 = \frac{-5f_1 + 4f_2 + f_3}{dx}
    \end{equation}

    The right boundary ($f^{\prime}_{n}$) is treated by utilizing
    the negative complex-conjugate of the Fourier image of the stencil
    at $f^{\prime}_1$:

    \begin{equation}
        f^{\prime}_{n} + 2f^{\prime}_{n-1s}
        =
        \frac{5f_{n} - 4f_{n-2} 1 f_{n-2}}{dx}
    \end{equation}

    The tridiagonal system that needs to be solved for the derivatives,
    then has the form:

    \begin{equation} \label{eqn:compact-tridiagonal-system}
     \begin{bmatrix}
         1&2\\
         1/4&1&1/4\\
         &1/4&1&1/4\\
         &&1/4&1&1/4\\
         &&&1/4&1&1/4\\
         &&&&&\ddots\\
         &&&&&&\ddots\\
         &&&&&&&\ddots\\
         &&&&&&&2&1
      \end{bmatrix}
      \begin{bmatrix}
          f^{\prime}_1 \\
          f^{\prime}_2 \\
          f^{\prime}_3 \\
          \vdots \\
          \vdots \\
          \vdots \\
          \vdots \\
          f^{\prime}_{n-1} \\
          f^{\prime}_n
       \end{bmatrix}
     =
     \begin{bmatrix}
         -5f_1 + 4f_2 + f_3\\
         f_{3} - f_{1}\\
         f_{4} - f_{2}\\
         \vdots\\
         \vdots\\
         \vdots\\
         \vdots\\
         f_{n} - f_{n-2}\\
         5f_{n} - 4f_{n-1} - f_{n-2}
      \end{bmatrix}
    \end{equation}

    A solution strategy is therefore concerned with the following:

    \begin{enumerate}
        \item{The evaluation of the right hand side.}
        \item{The solution of the resulting matrix system.}
    \end{enumerate}

\section{Reference approach}
    The current approach, developed for the CFDNS[ref] Compressible Navier-Stokes,
    and adapted to the Roadrunner[ref] architecture,
    is a parallel solver based on the standard LU algorithm,
    but specialized for the solution of a constant matrix.
    The problem is distributed among several ``processes'' and each
    process uses a single CPU core to perform computations.

    Assuming the coefficients of the LU algorithm are precomputed,
    the following steps are required to solve the tridiagonal system
    (only steps for the left-right sweep are listed, the right-left sweep
    consists of similar steps):

    \begin{enumerate}
    \item Precompute the arrays of coefficients of the LU algorithm---
    this is done once, in the beginning of the simulation process, and its cost
    can be neglected.
    \item Evaluate the RHS of the system.
    \item Compute $\hat{\phi_i}$ and $\hat{\psi_i}$ for each grid point $i$,
    with $\hat{\phi_i}$ and $\hat{\psi_i}$ both functions of the precomputed
    coefficient arrays and the right hand side.
    \item Every process posts the rightmost values of its $\hat{\phi_i}$ and $\hat{\psi_i}$,
        $\widetilde{\phi_k}$ and $\widetilde{\psi_k}$.
    \item Using the array of $\widetilde{\phi_k}$ and $\widetilde{\psi_k}$,
        each process computes $\widetilde{x_k}$, and uses it to compute
        $x_i,k = \phi_i + \psi_i\widetilde{x_k}$.
    \end{enumerate}

    The Message Passing Interface (MPI) is used for handling communication
    between processes.

\section{GPU architecture and OpenCL programming platform}

    \begin{figure}[h]
    \begin{center}
    \includegraphics[height=200pt]{img/tesla-block-diagram.jpg}
    \end{center}
    \caption{Distribution of a 2-D domain among processes}
    \label{fig:da-domain}
    \end{figure}

    GPUs afford massive parallelism by virtue of several (thousand)
    lightweight compute \emph{threads} or \emph{work-items}.
    The threads on a GPU are scheduled as \emph{blocks},
    which can be 1-D, 2-D, or 3-D.
    Each block is assigned to a \emph{streaming multiprocessor} (SM)
    on the GPU---the Tesla K20 GPU has 13 streaming multiprocessor---and
    all threads in a block are executed simultaneously by the SM.
    All SMs (and therefore all threads) have access to the GPU's ``global'' memory.
    Access to global memory is efficient when concurrent threads
    access neighbouring memory locations
    (known as \emph{coalesced memory access}).
    All threads in an SM also have access to a common fast, on-chip, \emph{shared memory}.
    Access to shared memory, in general, is much faster than access to global memory,
    but the amount of shared memory per SM is limited, typically a few dozen KB.
    Certain memory access patterns by concurrent threads to shared memory may result in
    serialization of the memory accesses, known as bank conflicts. These are detailed [here].

\section{Parallelization strategies}

    % left, lower, right, upper
    \begin{figure}[h]
    \begin{center}
    \includegraphics[trim={{100pt} {150pt} {100pt} {150pt}}, clip, height=200pt]{img/domain.eps}
    \end{center}
    \caption{Problem domain}
    \label{fig:domain}
    \end{figure}
    %
    \begin{figure}[ht]
    \begin{minipage}[b]{0.40\linewidth}
    \begin{center}
    \includegraphics[trim={{100pt} {150pt} {100pt} {150pt}}, clip, height=200pt]{img/one-block.eps}
    \end{center}
    \caption{A subdomain}
    \label{fig:block}
    \end{minipage}
    %
    \hspace{0.5cm}
    \begin{minipage}[b]{0.40\linewidth}
    \begin{center}
    \includegraphics[trim={{50pt} {150pt} {100pt} {150pt}}, clip, height=200pt]{img/one-thread.eps}
    \end{center}
    \caption{Single grid line within a subdomain}
    \label{fig:grid-line}
    \end{minipage}
    \end{figure}
    %
    \begin{figure}[h]
    \begin{center}
    \includegraphics[trim={{100pt} {250pt} {100pt} {150pt}}, clip, height=200pt]{img/block-dimensions.eps}
    \end{center}
    \caption{Domain}
    \label{fig:block-dimensions}
    \end{figure}

    \pagebreak

    The compact finite differences need to be evaluated
    at each grid point in the problem domain,
    shown in \ref{fig:domain}.
    Each grid point is represented as a cell in this representation.
    The domain has been shown as partitioned into subdomains \ref{fig:block}.
    The subdomains are three dimensional to facilitate
    the computation of derivatives in all three coordinate directions
    without incurring excessive communication costs.

    While evaluating the derivative in the x-direction (say),
    each subdomain is assigned to a \emph{compute unit}, or a \emph{process}
    which may be a single core, a multi-core processor, or a GPU.
    Each subdomain may be thought of as a collection of ``lines'' of grid points
    \ref{fig:grid-line}---referred to as \emph{grid lines}---in the x-direction,
    and the entire domain may be thought of as a collection of such subdomains.

    In the reference approach, each subdomain is assigned to a CPU core.
    Each core operates on a subdomain with dimensions as shown
    in the figure \ref{fig:block-dimensions}.
    The computational pattern followed by each core for the steps listed in Section[SEC]
    is a triple nested loop over the grid points in the subdomain,
    followed by a communication of the ``faces'' of the subdomain with the other
    cores:

    \begin{listing}
    \begin{minted}{fortran}

       do j=1,n2
          do i=1,n1
             x(1,i,j) = bet(1)*x(1,i,j)
             do k = 2, n3
                x(k,i,j) = (x(k,i,j) - a(k)*x(k-1,i,j))*bet(k)
             end do
             p_loc(i,j) = x(n3,i,j)
          end do
       end do

       call MPI_Allgather(p_loc,n1*n2,double_real,p,n1*n2, &
          double_real,comm,error)

    \end{minted}
    \caption{Computational pattern in reference approach}
    \label{listing:triple-loop}
    \end{listing}

    Conceptually, the innermost loop solves the tridiagonal system
    for a single grid line,
    while the outer loops navigate the different grid lines.
    The evalution of such triple loops are found to be the most time-consuming steps
    in the solution process,
    and it is desirable to introduce a second level of parallelism to replace these loops.
    Broadly speaking, there are two forms of parallelism that can be exploited:

    \textbf{Thread-level parallelism} \\
        In a multi-threaded environment,
        rather than having a single thread perform the local part
        of the computations for each line,
        it is beneficial to have several threads ``sweep'' through the subdomain,
        each performing the computations for a different tridiagonal system.
        This is a sort of \emph{thread parallel} operations (Figure \ref{fig:thread-parallel}).
        Algorithms that are inherently sequential,
        such as the Thomas algorithm lend themselves easily to thread-level parallelism
        With reference to the triple-loop in Listing 1,
        algorithms employing thread-level parallelism reduce the
        size of the outer two loops.

        \begin{figure}[h]
        \begin{center}
        \includegraphics[trim={{100pt} {150pt} {100pt} {150pt}}, clip, height=300pt]{img/thread-parallel.eps}
        \end{center}
        \caption{Parallel x}
        \label{fig:thread-parallel}
        \end{figure}

    \textbf{Data parallelism} \\
        Algorithms like Cyclic Reduction and Parallel Cyclic Reduction
        use parallel processors to perform
        perform the same operation at different points in the tridiagonal system.
        They are said to exhibit \emph{data parallelism} (Figure \ref{fig:data-parallel}).
        Although these algorithms perform a larger number of computations
        per step compared to the Thomas algorithm,
        they can take far fewer steps to run,
        depending on the number of available parallel processors.
        With reference to the triple-loop in Listing 1,
        algorithms employing data parallelism reduce the size of the innermost loop.

        \begin{figure}[h]
        \begin{center}
        \includegraphics[trim={{100pt} {150pt} {100pt} {150pt}}, clip, height=300pt]{img/data-parallel.eps}
        \end{center}
        \caption{Parallel yz}
        \label{fig:data-parallel}
        \end{figure}

    In this work, we evaluate the performance of a variant of Cyclic Reduction
    on the GPU that is both data-parallel and thread-parallel,
    and a thread-parallel implementation of the Thomas Algorithm on
    multi-core CPUs.

    Both implementations are compared against the reference implementation
    described in Section [SEC].
    The thread-parallel implementation of Thomas Algorithm is
    straightforward.
    However, implementation Cyclic Reduction on GPUs involves several issues,
    and our algorithm is described in detail in Section [SEC].
    %
    % \subsection{Evaluation of the RHS}
    %
    %     The RHS is evaluated by a 3 point stencil computation of
    %     the following form:
    %
    %     \begin{equation*}
    %         r_i = \frac{3}{4dx} (f_{i+1} - f_{i-1})
    %     \end{equation*}
    %
    %     This pointwise update can be performed easily in parallel
    %     for the internal points (points not close to the boundary of the process).
    %     Near the boundaries, the process will require information from its
    %     neighbours.
    %
    %     Thus, each process will require information from its
    %     neighbouring processes every time the derivative needs to be evaluated.
    %
    %     \begin{figure}[h]
    %     \begin{center}
    %     \includegraphics[trim={{100pt} {250pt} {100pt} {200pt}}, clip, height=200pt]{img/da-domain.eps}
    %     \end{center}
    %     \caption{Distribution of a 2-D domain among processes}
    %     \label{fig:da-domain}
    %     \end{figure}
    %
    %     \begin{figure}[h]
    %     \begin{center}
    %     \includegraphics[trim={{100pt} {250pt} {100pt} {200pt}}, clip, height=200pt]{img/da-local-portion.eps}
    %     \end{center}
    %     \caption{Elements required by a process to evaluate RHS}
    %     \label{fig:da-local-portion}
    %     \end{figure}
    %
    %     Thus, the RHS is evaluated in two steps:
    %
    %     \begin{enumerate}
    %         \item Exchange boundary information with neighbouring processes
    %         \item Evaluate the RHS at each grid point (in parallel)
    %     \end{enumerate}
    %
    %     The GPU is specialized for performing such a large number of independent
    %     computations as required in step 2.
    %     The RHS at each grid point is computing by a single work item
    %     (GPU thread).
    %     Therefore the number of work items scheduled is
    %     equal to the number of points in the local domain.
    %
    %     Another popular strategy is to take advantage of
    %     shared memory, by launching 2-dimensional blocks of work-items,
    %     loading each into shared memory, and streaming the third dimension
    %     in registers.
    %     This approach is discussed in [ref], and is much more effective
    %     for larger stencils---as each thread in a block will access the same
    %     value several times, it is effective to use shared memory as explicitly
    %     controlled cache.

\section{Description of the algorithm}

    The size of the problems solved using the high-order compact finite
    difference schemes discussed is usually much larger than can be accomodated
    on a single compute node.
    The problem domain is thus divided into ``subdomains'' and distributed among
    several processes.
    First, we discuss this cluster level parallelism.

    Given N equations in N variables of the form:

    \begin{align}
    & A =
     \begin{bmatrix}
         b_1 & c_1 \\
         a_2 & b_2 & c_2 \\
             & a_3 & b_3 & c_3 \\
             &     & a_4 & b_4 & c_4 \\
             &     &     &     &  \ddots & c_{N-1}\\
             &     &     &     &     a_N  & b_N
      \end{bmatrix}&
    \end{align}

    each process $p$ is concerned with the solution of a \emph{subsystem}
    of equations:

    \begin{align}
    &L_p =
    \begin{bmatrix}
        b_1^p & c_1^p \\
        a_2^p & b_2^p & c_2^p \\
              & a_3^p & b_3^p & c_3^p \\
              &       & a_4^p & b_4^p & c_4^p \\
              &       &       &       &  \ddots & c_{M-1}^p\\
              &       &       &       &     a_{M}^p  & b_{M}^p
     \end{bmatrix}&
    \end{align}

    For each subsystem, the following equations are solved:

    \begin{align}
        & L_p\boldsymbol{x}_p^R = \boldsymbol{r_p} &  \label{eqn:local-eqn-1} \\
        & L_p\boldsymbol{x}_p^{UH} = (-a_1^p, 0, 0 \hdots 0)^T &  \label{eqn:local-eqn-2} \\
        & L_p\boldsymbol{x}_p^{LH} = (0, 0, 0, \hdots -c_M^p)^T \label{eqn:local-eqn-3} &
    \end{align}

    The general solution is obtained as a linear combination of the above solutions:

    \begin{equation} \label{eqn:linear-combo}
        \boldsymbol{x}_p = {\xi}_p^{UH} \boldsymbol{x}_p^R + {\xi}_p^{LH} \boldsymbol{x}_p^{LH}
    \end{equation}

    Where the parameters ${\xi}_p^{UH}$ and ${\xi}_p^{LH}$ are obtained
    by solving the following ``reduced'' system:

    \begin{equation} \label{eqn:reduced-system}
     \begin{bmatrix}
         x_{1,M}^{LH}   &   -1                                                          \\
         -1             &   x_{2,1}^{UH} &   x_{2,1}^{LH}                               \\
                        &   x_{2,M}^{LH} &   x_{2,M}^{LH}  &  -1                        \\
         &              &   -1           &   x_{3,M}^{UH}  &  x_{1,M}^{LH}              \\
         &              &   &                x_{3,M}^{UH}  &  x_{1,M}^{LH} &    -1      \\
         &              &   &            &   &             \ddots   \\
         &              &   &            &   &             &        \ddots \\
         &  &   &   &   &   &   & -1 x_{P,1}^{UH}
      \end{bmatrix}
    \begin{bmatrix}
        \xi_1^{LH} \\
        \xi_2^{UH} \\
        \xi_2^{LH} \\
        \xi_3^{UH} \\
        \xi_3^{LH} \\
        \vdots \\
        \vdots \\
        \xi_P^{UH} \\
     \end{bmatrix}
     = -
     \begin{bmatrix}
         x_{1,M}^R \\
         x_{2,1}^R \\
         x_{2,M}^R \\
         x_{3,1}^R \\
         x_{3,M}^R \\
         \vdots \\
         \vdots \\
         x_P^R \\
      \end{bmatrix}
  \end{equation}

    The solution procedure is then:

    \begin{enumerate}
        \item Each process solves the three equations
        \ref{eqn:local-eqn-1}-\ref{eqn:local-eqn-3}.
        \item Each process communicates the parameters required to
        assemble the ``reduced'' system eqn:reduced-system
        \item The reduced system is solved for the coupling coefficients
        $\xi_i^{UH}$ and $xi_i^{LH}$, which are sent to their respective processes
        (if required)
        \item Each process solves \ref{eqn:linear-combo} for the local part
        of the solution to the global system.
    \end{enumerate}

    In the context of solving tridiagonal systems arising in compact finite differeces,
    we note the following:

    \begin{enumerate}
        \item Each process solves equations \ref{eqn:local-eqn-1}-\ref{eqn:local-eqn-3}
            for \emph{each grid line} in its subdomain.
        \item The left-hand side of these equations is the same
            for each grid line, coming from equation \ref{eqn:compact-tridiagonal-system}).
        \item The right-hand side for equations \ref{eqn:local-eqn-1} and
            \ref{eqn:local-eqn-2} are the same for each grid line.
            Thus, the equations need only be solved for a single grid line.
        \item Only \ref{eqn:local-eqn-3} is evaluated for each grid line
            in the subdomain. This is expected to be the most computationally expensive step.
    \end{enumerate}

    \subsection{Cyclic reduction}

        The classic Cyclic Reduction algorithm consists of two phases:
        forward reduction and backward substitution.
        The forward substitution phase begins with a system of size $N$,
        and by expressing every even-indexed equation $i$ as a linear
        combination of equations $i$, $i-1$ and $i+1$, reduces it to a
        system of size $N/2$.
        The process is repeated until a system of 2 equations in 2 unknowns
        is left.

        For each even-indexed $i$, we define:

        \begin{equation*}
        k_1 = \frac{a_i}{b_{i-1}},
        k_2 = \frac{c_i}{b_{i+1}}
        \end{equation*}

        Then, we update the values of $a$, $b$, $c$ and $d$ as:

        \begin{align} \label{eqn:forward-reduction}
        & a^{\prime}_i = -a_{i-1}k_1 & \\
        & b^{\prime}_i = b_i - c_{i-1}k_1 - a_{i+1}k_2 & \\
        & c^{\prime}_i = -c_{i+1}k_2 & \\
        & d^{\prime}_i = d_i - d_{i-1}k_1  - d_{i+1}k_2 &
        \end{align}

        The two equations are solved, yielding $x_N$ and $x_{N/2}$.
        In the backward substitution phase, every odd-indexed unknown $x_i$
        is solved for by substituting the known values of $x_{i-1}$ and $x_{i+1}$:

        \begin{align} \label{eqn:back-substitution}
        x_i = \frac{d^{\prime}_i - a^{\prime}_ix_{i-1} - c^{\prime}_ix_{i+1}}{b^{\prime}_i}
        \end{align}

        At each step, for the last index $i$, the forward reduction step is
        instead:
        \begin{align} \label{eqn:forward-reduction-right}
            & a^{\prime}_i = -a_{i-1}k_1 & \\
            & b^{\prime}_i = b_i - c_{i-1}k_1 & \\
            & d^{\prime}_i = d_i - d_{i-1}k_1 &
        \end{align}

        And the backward reduction step for $i=1$ is instead:
        \begin{align} \label{eqn:back-substitution-right}
        x_i = \frac{d^{\prime}_i - c^{\prime}_ix_{i+1}}{b^{\prime}_i}
        \end{align}

        \begin{figure}[h]
        \begin{center}
        \includegraphics[trim={{50pt} {300pt} {100pt} {50pt}}, clip, height=200pt]{img/forward-reduction.eps}
        \end{center}
        \caption{Cyclic reduction}
        \label{fig:forward-reduction}
        \end{figure}

        The cyclic reduction algorithm is well suited to the massively
        parallel architecture of the GPU.
        In the most straightforward implementation,
        we may launch a number of threads equal to
        half the number of equations,
        and assign each thread the task of
        solving the equations \ref{eqn:forward-reduction} for each even-indexed
        $i$.
        We can then perform the successive steps of forward reduction
        by halving the number of threads at each step.
        The backward substitution phase is completed in a similar way,
        with the number of threads doubling at each step.

        This naive approach leads to the following issues:

        \begin{enumerate}
            \item Most of the threads are inactive,
            except at the very beginning of forward reduction
            and the very end of backward substitution.

            \item The later steps of forward reduction and the first few
            steps of backward substitution involve memory accesses with
            large strides (see Figure \ref{fig:forward-reduction})
            If the threads are accessing global memory,
            this can lead to poor performance.
        \end{enumerate}

        The first issue is fairly straightforward to tackle.
        The OpenCL API allows \emph{scheduling} of a large number of
        threads---far more than the number of threads actually available
        on the computing device.
        Rather than scheduling threads for a single system (grid line),
        we may schedule threads to perform
        forward reduction steps for every grid line.

        To avoid non-coalesced accesses to global memory,
        we can instead have threads working with the fast, on-chip,
        \emph{shared memory}. In this case, the threads in each
        \emph{block} read data from global memory on to the block's
        shared memory.
        The threads then perform cyclic reduction on shared memory
        and write the data back to global memory.
        Conceptually, this approach has the advantage of avoiding
        non-coalesced global memory accesses, and facilitates data reuse.
        However, as described in [FastTridiagonalSolvers], the method
        suffers from so-called \emph{bank conflicts} in which,
        for certain memory access patters,
        data access by concurrent threads from shared memory may be serialized.
        This can seriously affect performance.
        Several approaches for avoiding bank conflicts have been proposed,
        this is at the cost of additional code complexity.
        Besides, an approach using shared memory suffers from the penalty
        of \emph{synchronization} between blocks and inactive threads.

    \subsection{Precomputing scheme}

        \begin{figure}[h]
        \begin{center}
        \includegraphics[trim={{50pt} {300pt} {100pt} {50pt}}, clip, height=200pt]{img/forward-reduction-coefficients.eps}
        \end{center}
        \caption{Forward reduction: indices with identical coefficients are shown in color}
        \label{fig:forward-reduction-coefficients}
        \end{figure}

        We introduce a variant of cyclic reduction,
        specialized for constant tridiagonal systems, that:

        \begin{enumerate}
        \item Reduces the number of accesses to global memory
        \item Reduces the memory requirement
        \item Reduces the number of computations in the forward reduction step
        \end{enumerate}

        We recall the form of the tridiagonal matrix:

        \[
        \begin{bmatrix}
            1&2\\
            1/4&1&1/4\\
            &1/4&1&1/4\\
            &&1/4&1&1/4\\
            &&&1/4&1&1/4\\
            &&&&&\ddots\\
            &&&&&&\ddots\\
            &&&&&&&\ddots\\
            &&&&&&&2&1
         \end{bmatrix}
         \]

        The matrix is the same for each grid line in the subdomain.
        Thus,
        the first three equations in \ref{eqn:forward-reduction}
        are the same for every grid line.
        For each forward reduction step,
        the parameters $a^{\prime}_i$, $b^{\prime}_i$ and $c^{\prime}_i$
        can be pre-computed.
        For a general tridiagonal system, this would require storage for
        $N/2 + N/4 + ... = 2N-2$ coefficients.
        However, it is readily apparent that,
        with the exception of the boundaries (near the first and last equations),
        the coefficients $k_1$, $k_2$,
        $a_i$, $b_i$ and $c_i$ are the same for each $i$.

        The size of the precomputed coefficient arrays is then $log_2(N)-1$
        (with an additional element for $a_N$ and $b_N$ as described below).
        For event modestly sized $N$ such as 512, this represents large savings
        in the memory requirements for the coefficient arrays.

        More importantly, at each forward reduction step,
        the number of non-coalesced memory accesses is greatly reduced.
        With the above precomputed coefficients,
        the forward reduction step now involves solving only a single equation
        for $d_i$ at each $i$.

        \begin{align} \label{eqn:forward-reduction-right}
            & a^{\prime}_i = -a_{i-1}k_1 & \\
            & b^{\prime}_i = b_i - c_{i-1}k_1 & \\
            & d^{\prime}_i = d_i - d_{i-1}k_1 &
        \end{align}

        The backward reduction step benefits from improved access to
        global memory.

        \subsubsection*{Boundary conditions}

        The tridiagonal systems for all grid lines are the same for all subdomains,
        except near the boundaries.
        From \ref{eqn:forward-reduction}, it is clear that
        to handle the boundaries, three additional sets of coefficients
        need to be stored:

        \begin{enumerate}
         \item The values for $b_1$ at each step
         \item The values for $k_1$ evaluated at $i=1$ for each step
         \item The values for $k_1$ evaluated at $i=n$ for each step,
            where $n$ is the last index of the reduction step.
        \end{enumerate}

        The values of $a_N$ and $b_N$ are required only for the
        final step (2-by-2 solve) of forward reduction.
        These two scalar values may be stored separately.

\section{Results and performance comparison}

    \subsection{Performance comparison}

    \subsection{Major concerns}

    \subsection{Performance improvement strategies}

\section{Computational practices}

\begin{figure}[h]
\begin{center}
\includegraphics[trim={{100pt} {150pt} {100pt} {150pt}}, clip, height=300pt]{img/data-and-thread-parallel.eps}
\end{center}
\caption{Parallel-both}
\label{fig:parallel-both}
\end{figure}

\end{document}