### 2015-11-14 thesis/acknowledgments.tex
delete junk files and reorganize

\chapter*{Acknowledgments}

I acknowledge everyone.