### 2015-11-14 thesis/introduction-and-background.tex

\chapter{Preliminaries}

\section{Graphics Processing Units (GPUs)}

\subsection{Motivation}

Most numerical approaches for modeling systems and phenomena
approximate continuous quantities as discrete quantities.
The larger the number of discrete entities used,
the better the approximation.
In CFD, the physical system (fluid)
is represented as a set of discrete points (finite difference methods),
or discrete cells (finite volume method),
and this discrete representation is termed as a \emph{mesh}.
In CFD applications
a mesh containing several points is used
to represent the underlying system,
and numerical simulations are performed
to study how physical quantities
change with time at these mesh points (or cells).
Typically, a \emph{time stepping} scheme is used to study this evolution,
wherein the simulation proceeds as a series of \emph{time steps}.
For each time step, the quantities at each mesh point at
\emph{updated} from quantities computed at the previous time step.

In the direct numerical simulation (DNS) of fluid systems,
the aim is to resolve the smallest features of the flow.
This requires the use of extremely fine meshes.
At each of the mesh points,
data about the flow must be stored
(for instance, the pressure, temperature and velocities at that point).
The large amount of data storage required places
restrictions on the problem sizes that can be accommodated
on a single workstation/PC.
Further,
the numerical simulation requires a large amount of
computations with this data.
In general, the computations are of a repetitive nature:
the same computations are performed repeatedly
for the different mesh points,
and for successive time steps.

Thus, CFD is typically a computationally expensive endeavour,
and it has long been of interest to use
\emph{parallel processing} to
perform numerical simulations of fluid flows.
This involves distributing computing tasks
among multiple independent processors,
which can run concurrently.
Of course, this requires that the computations be \emph{independent}
of each other.
Fortunately, the computational patterns that arise in CFD
typically satisfy this requirement:
the computations at different mesh points are largely independent
of each other and can be done in parallel.

The traditional approach to parallel computing has been
to distribute parallel tasks among CPU cores.
The cores may all have access to a common memory space
(shared memory systems),
or may have distinct memory spaces
(distributed memory systems).
Much more common than purely distributed systems are
\emph{hybrid} systems wherein the tasks may be distributed
among so-called \emph{nodes},
which are themselves shared-memory systems.
Again, the computational workhorse in all these systems has traditionally
been the CPU core.

A more recent parallel system that has seen
wide adoption
is the Graphics Processing Unit (GPU).
GPUs were developed primarily for performing
graphics computations, i.e.,
calculations with pixel data.
This is a \emph{highly} parallelizable task;
pixels can be processed independently and in parallel.
Thus, the GPU evolved as a system
to perform hundreds of computations concurrently,
and it should be no surprise that it soon
gained attention from the scientific computation community.
Now, several scientific software packages have been
written for GPUs,
and the worlds largest supercomputers employ GPUs
as accelerators.

\section{Compact finite differences}

Numerical evaluation of derivatives is a central component
in scientific computing.
The simplest and most widely-used approach for numerical differentiation
is the \emph{finite difference} approximation,
wherein the numerical approximation of the derivative
is expressed as a \emph{difference equation}.
A key application of the finite difference approximation
is in \emph{finite difference methods},
a family of numerical methods for solving differential equations
in which the derivatives are approximated using
finite difference approximations.
%
For example, we may consider a uniformly sampled
function $f(x)$,
sampled at points $x_1, x_2, x_3, \hdots x_n$.
At each sample point $i$, we may approximate the derivative as
a combination of the function values at
$i$, and its neighbouring points:

\begin{enumerate}
    \item When the derivative at $i$ is expressed
        as some combination of the function values
        at $i$, $i+1$, $i+2$, $\hdots$,
        we refer to the approximation as a
        \emph{forward} difference.

    \item When the derivative is expressed as
        some combination of the function values
        at $i$, $i-1$, $i-2$, $\hdots$,
        we refer to the approximation as a
        \emph{backward} difference.

    \item When the derivative is expressed as
        some combination of the function values
        on \emph{both} sides of $i$, i.e,
        at $\hdots$, $i-2$, $i-1$, $i$, $i-1$, $i-2$, $\hdots$,
        we refer to the approximation as a
        \emph{central} difference.
\end{enumerate}

The approximation of higher order derivatives generally requires
inclusion of a larger number of points in the
finite difference approximation
(referred to as the finite difference \emph{stencil}).
For a given order of derivative,
finite difference approximation of arbitrary stencil widths
may be derived,
with larger stencils associated with higher accuracy~\cite{fornberg1988generation}.
In most applications,
large stencil widths are undesirable.
This is because:

\begin{enumerate}
    \item The arithmetic intensity increases, i.e.,
        a larger number of computations must be performed per grid point.
    \item When the solution is done in parallel,
        the amount of boundary information that must be exchanged
        between processors increases.
\end{enumerate}

To overcome the problems associated with larger stencil widths,
another class of finite difference schemes,
known as \emph{compact} schemes may be used.
First proposed by Sanjeeva Lele ~\cite{lele1992compact},
compact schemes express the derivative at a point $i$
in terms of the derivatives at neighbouring points,
For example,
if $f_i$ represents the value of
a uniformly sampled function evaluated at the $i$th sample point,
the first derivative $f^{\prime}_i$ can be approximated from
a relation of the form:

\begin{equation}
\begin{split}
    \alpha(f^{\prime}_{i-1} + f^{\prime}_{i+1}) + \
        f^{\prime}_i
    \beta(f^{\prime}_{i-2} + f^{\prime}_{i+2}) + \
    \hdots
    = 
    a\frac{f_{i+1} - f_{i-1}}{dx} + \
    b\frac{f_{i+2} - f_{i-2}}{dx} + \\
    c\frac{f_{i+3} - f_{i-3}}{dx} + \
    \hdots
\end{split}
\label{eqn:general-compact}
\end{equation}
%
where $\alpha$, $\beta$, $a$, $b$, $c$, etc.,
are parameters that must satisfy certain constraints.
We note that for equations of the form in Eq. \ref{eqn:general-compact},
the derivative at any points $i$ cannot be computed explicitly.
Instead, we must write similar equations for \emph{all} points $i$
in the range.
This results in a system of linear equations
with unknowns \{$f^{\prime}_1$, $f^{\prime}_2$, $\hdots$, $f^{\prime}_n$\},
which may be represented by the matrix system $A\bm{x}=\bm{d}$,
where $\bm{x}$ is the vector
\{$f^{\prime}_1$, $f^{\prime}_2$, $\hdots$, $f^{\prime}_n$\},
$\bm{d}$ is a vector of right hand sides [Eq.\ref{eqn:general-compact}],
and $A$ is, in general, a \emph{banded} matrix.

The simplest classes of banded matrix include
diagonal, tridiagonal and pentadiagonal matrices.
When $A$ is diagonal ($\alpha$ = $\beta$ = $\hdots$ = 0),
the scheme is explicit and straightforward to evaluate:
the derivatives are simply computed pointwise
(in series or in parallel),
by application of the right-hand side stencil at each point.
When $A$ is tridiagonal ($\alpha \neq 0$, $\beta$ = $\hdots$ = 0),
the evaluation of the derivatives requires the solution
of the tridiagonal system, which yields the derivatives
at all points simultaneously.
The classical Thomas algorithm
also referred to as the Tridiagonal Matrix Algorithm,
is widely employed for this purpose.
The algorithm is straightforward to implement,
and stable for diagonally dominant matrices ~\cite{numericalrecipes}.