### 2015-11-15 thesis/compact-finite-difference.tex
update sections and add content to thesis

\chapter{Application to compact finite difference evaluation}

\section{Compact finite difference evaluation strategy}
    \subsection{Shared memory systems}
    \subsection{Distributed memory systems}
    \subsection{Heterogeneous systems}

\section{Distributed tridiagonal solver}
    \subsection{General algorithm}
    \subsection{Specialization for compact finite difference
        evaluations}

\section{GPU application}
    \subsection{General strategy}
    \subsection{GPU-GPU communication}
        \subsubsection{Naive strategy}
        \subsubsection{NVIDIA GPUDirect Technology}
    \subsection{Implementation}
        \subsubsection{Precomputing coefficients}
        \subsubsection{Solving secondary systems}
        \subsubsection{Computing right hand sides}
        \subsubsection{Solving primary systems}
        \subsubsection{Constructing and solving reduced systems}
        \subsubsection{Summing primary and secondary solutions}
        \subsubsection{Evaluating derivatives in other
            coordinate directions}