### 2015-11-15 thesis/introduction-and-background.tex

\chapter{Introduction and Background}

\section{Motivation}

In the direct numerical simulation (DNS) of fluid systems,
the aim is to resolve the smallest features of the flow
without resorting to any modeling of the turbulence.
This involves the use of
extremely fine computational grids to
discretize the flow domain,
and some numerical method for solving the
flow equations on this grid.
By far the most popular numerical approach for
solving the flow equations
is the \emph{finite difference method}.
The finite difference method approximates the
spatial and temporal derivatives appearing in the
partial differential equations that describe the flow
using finite difference schemes.
\emph{Compact finite difference schemes} are a class
of finite difference schemes that have found widespread
adoption in DNS codes
for their high order of accuracy.
The evaluation of compact finite differences
requires the repeated solution of \emph{banded} linear systems,
making them fairly complex and expensive computationally.

As mentioned, DNS involves the use of extremely fine grids.
At each of the grid points,
various data about the flow must be stored---for
example, the geometric coordinates (x, y and z),
pressure, temperature and velocities at that point.
The amount of memory required to store this data
for even small problem sizes
can quickly exceed the capacity of modern workstations/PCs.
Thus, a \emph{distributed memory} parallel system
is generally required for performing DNS.
Here, the traditional approach has been
to distribute parallel tasks among individual CPU cores,
or groups of CPU cores
that share common memory spaces.
In the latter case,
each group of cores constitutesa \emph{shared memory} system,
and the overall system is referred to as a \emph{hybrid system}.

The workhorse for computation in the above described
parallel systems is the CPU core,
however, more recently,
Graphics Processing Units (GPUs)
are being used for performing intensive calculations.
GPUs, while themselves being highly parallel processors,
can also function as accelerators in distributed memory systems
(GPU clusters).
However,
applications that exploit such systems
need careful redesign of algorithms---and
sometimes, substantial changes to code---to
see significant performance improvements.
This is because the CPU and GPU have very different architectures,
and any na\"{\i}ve ``parallelization'' of algorithms
designed for the CPU
is likely not take full advantage of the GPU's
memory heirarchy and compute ability.

The objective of this work is to develop an
approach for evaluating
compact finite differences
on multiple GPUs in a distributed system.
This is of interest because
the evaluation of spatial derivatives using
compact finite difference schemes is one of the most
expensive tasks in DNS.
But perhaps more significantly,
it encompasses several computational patterns
such as pointwise updates, stencil evaluations,
and solutions of distributed tridiagonal systems.
Efficient implementation of these computational patterns
on GPUs is of interest in other areas of CFD,
and scientific computation in general.

\section{Compact finite differences}

Numerical evaluation of derivatives is a central component
in scientific computing.
The simplest and most widely-used approach for numerical differentiation
is the \emph{finite difference} approximation,
wherein the numerical approximation of the derivative
is expressed as a \emph{difference equation}.
A key application of the finite difference approximation
is in \emph{finite difference methods},
a family of numerical methods for solving differential equations
in which the derivatives are approximated using
finite difference approximations.
%
For example, we may consider a uniformly sampled
function $f(x)$,
sampled at points $x_1, x_2, x_3, \hdots x_n$.
At each sample point $i$, we may approximate the derivative as
a combination of the function values at
$i$, and its neighbouring points:

\begin{enumerate}
    \item When the derivative at $i$ is expressed
        as some combination of the function values
        at $i$, $i+1$, $i+2$, $\hdots$,
        we refer to the approximation as a
        \emph{forward} difference.

    \item When the derivative is expressed as
        some combination of the function values
        at $i$, $i-1$, $i-2$, $\hdots$,
        we refer to the approximation as a
        \emph{backward} difference.

    \item When the derivative is expressed as
        some combination of the function values
        on \emph{both} sides of $i$, i.e,
        at $\hdots$, $i-2$, $i-1$, $i$, $i-1$, $i-2$, $\hdots$,
        we refer to the approximation as a
        \emph{central} difference.
\end{enumerate}

The approximation of higher order derivatives generally requires
inclusion of a larger number of points in the
finite difference approximation
(referred to as the finite difference \emph{stencil}).
For a given order of derivative,
finite difference approximation of arbitrary stencil widths
may be derived,
with larger stencils associated with higher accuracy~\cite{fornberg1988generation}.

\subsection{Wavenumber analysis of finite difference methods}

For DNS applications, the relevant measure of accuracy of a
finite difference scheme is obtained from the so-called
\emph{modified wavenumber} approach.
Here, we test the scheme's ability to
accurately estimate the derivative of sinusoidal functions
with increasing wavenumbers (frequencies),
given a specified grid size. 
We expect that the approximation of the derivative
becomes more difficult with increasing wavenumbers,
as the function value varies more rapidly.
The wavenumber of the approximate derivative
as given by the finite difference scheme is compared with
the wavenumber of the exact derivative.
The result for different schemes is shown
in Fig. \ref{fig:modified-wavenumbers}.

\begin{figure}
\begin{center}
\includegraphics[height=250pt]{fig/modified-wavenumbers.eps}
\end{center}
\caption{Modified wavenumbers for three different schemes
    with the same stencil width.
    The compact schemes better estimate the derivative
    for higher wavenumbers.}
\label{fig:modified-wavenumbers}
\end{figure}

The ability of a finite difference scheme
to accomodate large wavenumbers
is extremely relevant in
computational fluid dynamics applications
~\cite{KravchenkoEffNumErr}.
In DNS, the scheme
\emph{must} capture the
rapidly varying characteristics of the flow
associated with the turbulence.
For this purpose,
higher order explicit schemes
may be considered.
While they are traightforward to compute,
they are associated with large stencil widths.
In most applications,
large stencil widths are undesirable.
This is because
the arithmetic intensity increases with stencil size, i.e.,
a larger number of computations must be performed per grid point.
Further,
the amount of boundary information that must be exchanged
between parallel processes increases,
which can seriously affect overall performance.

From \ref{fig:modified-wavenumbers},
it is clear that the compact schemes are better
able to compute derivatives for larger wavenumbers
for a given stencil width.
Thus, they are widely used in
CFD for evaluation of spatial derivatives.

\subsection{General form of compact finite difference schemes}
\label{subsec:gen-form-compact}

Compact schemes express the derivative at a point $i$
in terms of
function values \emph{and derivatives}
at neighbouring points,
For example,
if $f_i$ represents the value of
a uniformly sampled function evaluated at the $i$th sample point,
the first derivative $f^{\prime}_i$ can be approximated from
a relation of the form:
%
\begin{equation}
\begin{split}
    \alpha(f^{\prime}_{i-1} + f^{\prime}_{i+1}) + \
        f^{\prime}_i
    \beta(f^{\prime}_{i-2} + f^{\prime}_{i+2}) + \
    \hdots
    = 
    a\frac{f_{i+1} - f_{i-1}}{dx} + \
    b\frac{f_{i+2} - f_{i-2}}{dx} + \\
    c\frac{f_{i+3} - f_{i-3}}{dx} + \
    \hdots
\end{split}
\label{eqn:general-compact}
\end{equation}
%
where $\alpha$, $\beta$, $a$, $b$, $c$, etc.,
are parameters that must satisfy certain constraints ~\cite{lele1992compact}.
We note that for equations of the form in Eq. \ref{eqn:general-compact},
the derivative at any points $i$ cannot be computed explicitly.
Instead, we must write similar equations for \emph{all} points $i$
in the range.
This results in a system of linear equations
with unknowns \{$f^{\prime}_1$, $f^{\prime}_2$, $\hdots$, $f^{\prime}_n$\},
which may be represented by the matrix system $A\bm{x}=\bm{d}$,
where $\bm{x}$ is the vector
\{$f^{\prime}_1$, $f^{\prime}_2$, $\hdots$, $f^{\prime}_n$\},
$\bm{d}$ is a vector of right hand sides [Eq.\ref{eqn:general-compact}],
and $A$ is, in general, a \emph{banded} matrix.

\subsection{Boundary conditions}

One of the advantages of the compact finite difference approach
is that it accommodates \emph{non-periodic} boundary conditions.
This is in contrast to other methods used in DNS,
such as the \emph{spectral} methods.
We note that the Eq. (\ref{eqn:general-compact})
cannot be applied near the boundary points.
At the boundaries,
\emph{non-centered} or \emph{one-sided}
finite difference approximations are required.
Some considerations are made in choosing
these approximations:
firstly, the bandwidth of the resulting banded matrix
must be preserved.
Secondly,
the width of the boundary stencils
must be lower than the interior stencils,
as higher order boundary stencils are unstable
~\cite{kennedy1994several}.
In general, boundary equations for the first derivative
are of the following form ~\cite{lele1992compact}:

\begin{equation}
    f_1^{\prime} + \alpha f_2^{\prime} = \
        \frac{1}{h}(af_1 + bf_2 + cf_3 + df_4)
\label{eqn:boundary-compact}
\end{equation}

\subsection{Tridiagonal compact schemes}

As mentioned in Sec. \ref{subsec:gen-form-compact},
compact finite difference schemes lead to
\emph{banded} linear systems.
The simplest classes of banded matrix include
diagonal, tridiagonal and pentadiagonal matrices.
When $\alpha$ = $\beta$ = $\hdots$ = 0,
$A$ is a $\emph{diagonal}$ matrix.
In this case, the
scheme is \emph{explicit}
and the derivatives are straightforward to evaluate:
this simply involves the
application of the right-hand side stencil at each point.
When $\alpha \neq 0, \beta = \hdots$ = 0,
$A$ is \emph{tridiagonal}.
The evaluation of the derivatives requires
the solution of the resulting tridiagonal system.
The classical Thomas algorithm---also
referred to as the Tridiagonal Matrix Algorithm---is
widely employed for this purpose.
The algorithm is straightforward to implement,
and stable for diagonally dominant matrices ~\cite{numericalrecipes}.
When $\alpha \neq 0$ and $\beta \neq  0$,
pentadiagonal and higher-bandwidth systems arise.
These can be more expensive and difficult to evaluate
than tridiagonal systems
using a direct method.

This work will focus on
compact finite difference schemes that lead
to tridiagonal systems.
An example of such a scheme is obtained
by substituting 
$\beta = 0$, $\alpha = \frac{1}{4}$ and $a = \frac{3}{4}$
in Eq. \ref{eqn:general-compact}
(all other coefficients are set to zero).
This leads to a fourth-order accurate
tridiagonal compact finite difference scheme,
known also as the Pad\'{e} scheme.
At the boundaries,
the following \emph{third-order} accurate
approximations are used:

\begin{align}
    f^{\prime}_1 + 2f^{\prime}_2 = \frac{-5f_1 + 4f_2 + f_3}{dx} \\
    f^{\prime}_{n} + 2f^{\prime}_{n-1}
    =
    \frac{5f_{n} - 4f_{n-2} -  f_{n-1}}{dx}
\end{align}
%
The resulting tridiagonal system is then:

\begin{equation}
 \label{eqn:compact-tridiagonal-system}
 \begin{bmatrix}
     1&2\\
     1/4&1&1/4\\
     &1/4&1&1/4\\
     &&1/4&1&1/4\\
     &&&1/4&1&1/4\\
     &&&&&\ddots\\
     &&&&&&\ddots\\
     &&&&&&&\ddots\\
     &&&&&&&2&1
  \end{bmatrix}
  \begin{bmatrix}
      f^{\prime}_1 \\

      f^{\prime}_2 \\
      f^{\prime}_3 \\
      \vdots \\
      \vdots \\
      \vdots \\
      \vdots \\
      f^{\prime}_{n-1} \\
      f^{\prime}_n
   \end{bmatrix}
 =
 \begin{bmatrix}
     \frac{-5f_1 + 4f_2 + f_3}{2}\\
     \frac{3(f_{3} - f_{1})}{4}\\
     \frac{3(f_{4} - f_{2})}{4}\\
     \vdots\\
     \vdots\\
     \vdots\\
     \vdots\\
     \frac{3(f_{n} - f_{n-2})}{4}\\
     \frac{5f_{n} - 4f_{n-1} - f_{n-2}}{2}
  \end{bmatrix}
\end{equation}

Similar tridiagonal schemes are available
to evaluate higher order derivatives.
Compact schemes for
second, third and fourth derivatives
are derived in ~\cite{lele1992compact}.

\section{Graphics processing units}

\subsection{Introduction}
\subsection{Memory model}
\subsubsection{Global memory}
\subsubsection{Shared memory}
\subsubsection{Registers and local memory}
\subsection{CUDA Programming model}
\subsubsection{Kernels}
\subsubsection{Thread organization}

\section{Tridiagonal solvers}
\subsection{Thomas algorithm}
\subsection{Cyclic reduction}
\subsection{Other algorithms}