### 2015-11-15 thesis/proposed-tridiagonal-algorithm.tex

\chapter{Proposed tridiagonal algorithm}

\section{Modified cyclic reduction}
\subsection{Analysis of forward reduction coefficiets}
\subsection{Precomputing coefficients}
\subsection{Forward reduction}
\subsection{Back substitution}

\section{Implementation}
\subsection{General strategy}
\subsection{Global memory implementation}
\subsubsection{Algorithm}
\subsubsection{Kernels}
\subsubsection{Grid and thread block configuration}

\subsection{Shared memory implementation}
\subsubsection{Algorithm}
\subsubsection{Kernel}
\subsubsection{Grid and thread block configuration}