### 2015-11-15 thesis/results.tex

\chapter{Results}

\section{Validation}
    \subsection{Validation of tridiagonal solver}
    \subsection{Validation of compact finite difference application}
        \subsubsection{Tests for individual kernels}
        \subsubsection{Tests for communication routines}
        \subsubsection{Test functions for derivatives}
        \subsubsection{Order of convergence tests}
    \subsubsection{Validation results}

\section{Performance of GPU tridiagonal solver}
\section{Performance of compact finite difference application}