### 2015-11-17 thesis/abstract.tex
add chapter on near-toeplitz solver

\chapter*{Abstract}

We study the impact of the matrix structure on
the cyclic reduction algorithm
and show that that precomputing
forward reduction coefficients can be
an especially effective strategy.
his is the abstract