### 2015-11-17 thesis/acknowledgments.tex

\chapter*{Acknowledgments}

I owe my deepest thanks
to my advisor, Dr. Richard S. Miller,
for hist immense guidance and support;
and the members of my committee,
Dr. Gang Li and Dr. Lonny Thompson
for their valuable time.
I would like to acknowledge
Dr. Daniel Livescu at Los Alamos National Lab
for introducing us to this interesting problem,
for his helpful comments,
and for generous funding.
I also thank
Dr. Melissa Smith and Karan Sapra
for their helpful guidance,
and for organizing an excellent course on GPU computing.
I owe a great deal of gratitude to
Dr. Christopher Cox at Clemson University
for his guidance on a range of topics
and for many interesting conversations.
I would like to thank the
staff of the Cyberinfrastructure Technology Integration,
and especially
Dr. Galen Collier and Dr. Marcin Ziolkowski
for their valuable comments and support.
I owe thanks to Clemson University for the generous
allocation of compute time on the Palmetto cluster.
I acknowledge
Dr. K. N. Seetharamu and Dr. V. Krishna,
my advisors at PES Institute of Technology
for inspiring in me a passion for
scientific research.
I owe a special thanks to the
Software Carpentry Foundation
and its members,
for helping me be a
a more productive programmer,
a more deliberate learner,
a more effective teacher,
and a better person.

Finally,
I would like to thank my family and friends
for everything they have done for me.