### 2015-11-17 thesis/conclusions.tex

\chapter{Conclusions and Future Work}

We have presented an efficient approach for evaluating
compact finite differences on GPU-accelerated clusters.

The integration of this approach to our
current DNS codes is yet to be performed.
To ameliorate the cost of host-device transfers,
it is likely that other portions of the code
will also need to be ported to GPUs.
However,
most of the computational patterns
followed in the rest of the code
have been covered in the current work,
and fewer challenges are likely to be faced
in this process.

At the core of our approach is a
fast tridiagonal solver for the resulting linear systems.
Here, we make use of the simple matrix structure
to obtain better performance.
The applicability of this strategy to the
parallel cyclic reduction (PCR) algorithm,
and to other hybrid approaches is yet to be studied.

Strategies for the evaluation of compact finite differences
on other parallel architectures is of great interest,
especially for the Intel Many Integrated Core (MIC)
architectures.