### 2015-11-18 thesis/abstract.tex
corrections - round 1

\chapter*{Abstract}

\emph{Compact} finite difference schemes are widely used
in the direct numerical simulation of fluid flows
for their ability to better resolve the small scales of turbulence.
However, they can be expensive to evaluate and
difficult to parallelize.
In this work, we present an approach for the computation
of compact finite differences
and similar tridiagonal schemes
on graphics processing units (GPUs).
We present a variant of the \emph{cyclic reduction} 
algorithm for solving the
tridiagonal linear systems that arise in such numerical schemes.
We study the impact of the matrix structure on
the cyclic reduction algorithm
and show that precomputing
forward reduction coefficients can be
especially effective for obtaining good performance.
Our tridiagonal solver is able to outperform
the NVIDIA CUSPARSE
and the multithreaded Intel MKL tridiagonal solvers
on GPU and CPU respectively.
In addition,
we present a parallelization strategy for GPU-accelerated clusters,
and show scalabality
of a 3-D compact finite difference application
for up to 64 GPUs on Clemson's Palmetto cluster.