### 2015-11-30 presentation/results.tex

\section{Results}

\begin{frame}
\frametitle{Overview of results}
\begin{itemize}
\item Results for tridiagonal solver
\begin{itemize}
    \item Global memory v/s shared memory implementations
    \item Compared against library solvers
        (multicore CPU and GPU)
\end{itemize}
\item Results for compact finite difference application
\begin{itemize}
    \item Profiling results
    \item Scaling results
    \item Comparison with CPU-only approach
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Global memory v/s shared memory implementations}
\begin{columns}
\begin{column}{0.5\textwidth}
\begin{itemize}
\item For 2-D problems, shared memory clearly better
\item For 3-D problems, shared memory speedup
    is smaller
\item Global memory makes better use of
    GPU parallelism in 3-D case
\end{itemize}
\end{column}
\begin{column}{0.5\textwidth}
\centering
\includegraphics[width=140px]{img/global-vs-shared-2d.eps}

\includegraphics[width=140px]{img/global-vs-shared-3d.eps}
\end{column}
\end{columns}
\end{frame}

\begin{frame}
\frametitle{Comparison with library solvers - timings}
\begin{table}
\resizebox{\textwidth}{!}{%
\input{table/bench.tex}
}
\end{table}
\end{frame}

\begin{frame}
\frametitle{Comparison with library solvers - speedup}
\centering
\includegraphics[width=180px]{img/bench-2d.eps}
\includegraphics[width=180px]{img/bench-3d.eps}
\begin{table}
\end{table}
\end{frame}

\begin{frame}
\frametitle{Note on speedups}
\begin{itemize}
\item Comparing against 8 CPU cores, speedups in the range 1-2x
\item Solving the tridiagonal systems
    is the \emph{least amenable to efficient parallel solution}
\item Objective is to keep the problem on the GPU
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Compact finite difference evaluation - profiling}
\begin{columns}
\begin{column}{0.5\textwidth}
\begin{itemize}
    \item Problem sizes: $1024^3$ and $2048^3$ on 64 GPUs
    \item For larger problems, more time spent
        on tridiagonal systems
    \item Significant portion of time spent on
        data permutation
\end{itemize}
\end{column}
\begin{column}{0.5\textwidth}
\centering
\includegraphics[width=170px]{img/profiling-1024-64.eps}

\includegraphics[width=170px]{img/profiling-2048-64.eps}
\end{column}
\end{columns}
\end{frame}

\begin{frame}
\frametitle{Compact finite difference evaluation - strong scaling}
\centering
\includegraphics[width=180px]{img/strong-scaling-256.eps}
\includegraphics[width=180px]{img/strong-scaling-512.eps}

Problem sizes: $256^3$ and $512^3$
\end{frame}

\begin{frame}
\frametitle{Compact finite difference evaluation - weak scaling}
\centering
\includegraphics[width=180px]{img/weak-scaling-128.eps}
\includegraphics[width=180px]{img/weak-scaling-256.eps}

Problem sizes: $128^3$ and $256^3$ \emph{per process}
\end{frame}

\begin{frame}
\frametitle{Comparison with CPU-only approach}
\begin{columns}
\begin{column}{0.5\textwidth}
\begin{itemize}
\item Reference implementation: approach in CFDNS code
    (Los Alamos National Lab)
\item Maintain a ratio of 8 CPU cores : 1 GPU in comparison
\end{itemize}
\end{column}
\begin{column}{0.5\textwidth}
\centering
\begin{table}
\resizebox{0.75\textwidth}{!}{%
\input{table/compact-refimpl-timings.tex}
}
\end{table}

\includegraphics[width=140px]{img/compact-refimpl-speedups.eps}
\end{column}
\end{columns}
\end{frame}